---
title: "Virtual species paper analyses"
author: "Jan Salecker"
date: "8/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The spectre manuscript

The spectre manuscript will probably be focused around the description of the algorithm and the R package.
Thus, a technical usecase is need for several reasons:

* To provide an example how the algorithm can be used and how the inputs and outputs look like
* To assess the efficiency of the spectre algorithm under various scenarios
* To point out future directions of package development

In addition to Matthias (@mspngbrg) approach (investigating the BCI dataset), this draft uses the virtualspecies data generation approach (see `virtualspecies_01_simple.Rmd` and `virtualspecies_02_advanced.RMD`) to explore the spectre package.


## Executing


```{r packages}
library(spectre)
library(tidyverse)
library(future)
library(virtualspecies)
library(furrr)
library(clustermq)
library(landscapetools)

## Source spectre analysis functions (R folder)
sapply(list.files("../R", full.names = TRUE), source)
set.seed(3562347)
exec <- "HPC" # "HPC"
dir.hpc <- file.path("/home/uni08/jsaleck/spectre")
dir.cloud <- file.path("/home/jan/netlogo_jan/spectre")
dir.local <- file.path("spectre/vignettes")

## Constant parameters:
replicates <- 20
max_runs <- 100000
energy_threshold <- 0
beta <- 0.75

## Variable parameters:
landscape_size <- c(10,20,30)
corr_within <- c(0,0.5,1)
corr_among <- c(0,0.05,0.1)
gamma <- c(100,200,300)
random_seeds <- round(runif(replicates) * 100000)

## Generate parameter matrix:
parameters <- expand.grid(landscape_size = landscape_size, 
                          corr_within = corr_within,
                          gamma = gamma, 
                          beta = beta,
                          corr_among = corr_among,
                          random_seeds = random_seeds)

if (exec == "local") {
  plan(multisession)
  results <- furrr::future_map_dfr(seq(nrow(parameters)), function(x) {
    virtualspecies_simfun(siminputrow = x, 
                          parameters = parameters,
                          max_runs = max_runs,
                          energy_threshold = energy_threshold,
                          writeRDS = FALSE,
                          outdir = dir.cloud)
  })
}
if (exec == "HPC") {
  
  maxjobs.hpc <- 2000
  njobs <- min(nrow(parameters), maxjobs.hpc)
  jobIDs <- seq(nrow(parameters))
  
  results <- clustermq::Q(fun = virtualspecies_simfun, 
                          siminputrow = jobIDs,
                          const = list(parameters = parameters,
                                       max_runs = max_runs,
                                       energy_threshold = energy_threshold,
                                       writeRDS = TRUE,
                                       outdir = dir.hpc),
                          export = list(), 
                          seed = 42, 
                          n_jobs = njobs, 
                          template = list(job_name = "spectre", # define jobname
                                          log_file = "spectre.log", # define logfile name
                                          queue = "medium",  # define HPC queue
                                          service = "normal", # define HPC service
                                          walltime = "24:00:00", # define walltime
                                          mem_cpu = "4000")) # define memory per cpu   
  results <- dplyr::bind_rows(results)
}

#### Restore results from rds files if neccessary:
results <- purrr::map_dfr(list.files(dir.cloud, pattern = "rds", full.names = TRUE), function(x) {
  res.x <- readRDS(x)
  return(res.x)
})

#### STORE RESULTS:
saveRDS(results, file=file.path(dir.cloud, "virtualspecies_03_manuscript_data.rds"))

```

## Analysis


#### Energy over time


#### Minimum energy

standardized regression coefficient:
summary(lm(scale(energy) ~ scale(landscape_size) * scale(gamma) * scale(corr_within) * scale(corr_among), data=energy_min))

#### Realtive commonness error



